Confirmation of \srcfull{} as a long period radio transient

### 2025-05-23 paper/main.tex
Various updates from co-author suggestions

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. J. McSweeney et al.]{
Samuel J. McSweeney$^{1}$,
Natasha Hurley-Walker$^{1}$,
Cs\'{a}nad Horv\'{a}th$^{1}$,
Akash Anumarlapudi$^{2,3}$,
\newauthor
Angie Waszewski$^{1}$,
Dougal Dobie$^{2,3}$,
John Morgan$^{1}$,
David Kaplan$^{?}$,
and Ziteng Wang$^{1}$
\\
% List of institutions
$^{1}$International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia\\
$^{2}$Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia\\
$^{3}$ARC Centre of Excellence for Gravitational Wave Discovery (OzGrav), Hawthorn, VIC 3122, Australia
}

@@

\src{} was detected blindly with the Murchison Widefield Array \citep[MWA;][]{Tingay2013} in observations taken between June--September 2024 under the `Galactic Plane Monitor', project code G0080. The prime objective of the campaign was to search for long-period transients by scanning the Galactic plane over $285^\circ < l < 65^\circ$, $|b|<\pm15^\circ$ on a bi-weekly cadence at 185--215\,MHz \citep[see Methods of][]{2023Natur.619..487H}; a full description will be released by Hurley-Walker et al., in prep. We also observed over June--September 2022 but made no detections in these data, as reported in \citetalias{2024MNRAS.535..909D}.
Subsequent to its redetection, we commenced a Director's Discretionary Time (DDT) campaign under project code D0042, to obtain further observations, adjusting the frequency from 200.32\,MHz to 184.96\,MHz for some of the observations to attempt to improve measurements of its DM.
Time of arrival (ToA) predictions for the DDT observations were made using a preliminary ephemeris derived from the GPM detections, which, by virtue at being at the same or similar frequencies, absorbed the delays due to dispersion and scattering into the `PEPOCH' term.

We imaged both the 2024 GPM data and the follow-up DDT data at the location of \src{} using \citep[\textsc{WSClean}][]{2014MNRAS.444..606O}, masking \src{} itself, forming a deep model of the sky for each observation. After subtracting this from the visibilities, we phase-rotated to \src{}, and averaged the baseline data to produce a dynamic spectrum.

@@

Paper~1 presented an exhaustive search of the ASKAP data available at the time of writing, comprising 60\,h of data, and yielding only a single detection of \src{} (in Scheduling Block ID 47253). Subsequent to the writing of Paper~1, a further epoch of the VAST survey was performed (SBID 63600) in which \src{} produced a detectable pulsation.

@@

The correlator data were calibrated using the standard SARAO SDP calibration pipeline, and imaged using \textsc{WSClean}. We formed a deep model of the sources within the primary beam, excepting \src{}, and subtracted this from the visibilities. We then averaged the visibilities, excluding all baselines shorter than 150\,m, to produce a dynamic spectrum of the source, shown in \Fig~\ref{fig:scattering_DM}.

We also re-imaged the original 2023-05-28 L-band MeerKAT observation described in \citetalias{2024MNRAS.535..909D}, which was taken under project code DDT-20230525-DD-01 under capture block 1685306788. We performed a similar processing of the data, and identified four pulses with ToAs consistent with an ephemeris derived from the other data. The folded and stacked dynamic spectra are shown in \Fig~\ref{fig:scattering_DM}.

@@

The low-frequency asymmetry is consistent with a scattering time scale of a few tens of seconds at MWA frequencies, which scales to a few tens of milliseconds at 1\,GHz.
This is in broad agreement with the model predictions of NE2001 ($\tau_{\rm sc,1GHz} = 66\,$ms) and YMW16 ($\tau_{\rm sc,1 GHz} = 75\,$ms) at the source's location assuming the DM value of 710\,pc\,cm$^{-3}$ reported in \citepalias{2024MNRAS.535..909D}.

To account for the asymmetry, we fit exponentially modified Gaussians to each pulse with the form,
\begin{equation}
  \emg(t) = A \frac{\sigma}{\tau}\sqrt{\frac{\pi}{2}}
    \exp\left(-\frac12 \left( \frac{t - \mu}{\tau} \right)^2 \right)  \erfcx \left(\frac{1}{\sqrt{2}} \left(\frac{\sigma}{\tau} - \frac{t - \mu}{\tau} \right) \right),
  \label{eqn:emg}
\end{equation}
where $\erfcx(X) = \exp(X^2) \erfc(X)$ is the scaled complimentary error function; $A$, $\mu$, and $\sigma$ are the height, location, and scale of an unscattered Gaussian pulse; and the scattering time scale was fixed at
\begin{equation}
      \tau(\nu) = 0.07\, \left(\frac{\nu}{\rm GHz}\right)^{-4} \, {\rm s},
\end{equation}
according to each observation's centre frequency.
A pulse stack containing all observed pulses and the corresponding unscattered pulses derived from the fits is shown in \Fig~\ref{fig:pulsestack}.

\begin{figure*}
      \centering
          \includegraphics[width=0.95\linewidth]{pulsestack.pdf}
              \caption{Pulsestack of barycentred, dedispersed profiles from the ephemeris. The dashed red vertical line marks zero phase. Baselines have been fitted and subtracted from each lightcurve. The blue dashed lines show (normalised) fitted Gaussians without scattering (see main text for details).}
                  \label{fig:pulsestack}
\end{figure*}

@@

The apparent lack of linear polarisation in the MWA pulses, however, is expected due to the high RM of ${\sim}961\,{\rm rad}/{\rm m}^2$ \citepalias[as reported in][]{2024MNRAS.535..909D}, causing the pulses to be depolarised within individual channels.
We also report that no significantly circularly polarised emission was seen at MWA frequencies.

@@

The high S/N of each time bin allowed us to test whether the rotation measure (RM) was consistent across each pulse.
We therefore performed a joint fit of the Stokes Q and U spectra of each time bin to
\begin{equation}
    Q(\nu) = {\rm Re}(\mathcal{L}(\nu)),
    \quad
    U(\nu) = {\rm Im}(\mathcal{L}(\nu)),
    \label{eqn:RM}
\end{equation}
where
\begin{equation}
    \mathcal{L}(\nu)
        = L_{\rm 1\,GHz}\left(\frac{\nu}{\rm 1\,GHz}\right)^{\alpha_L} e^{2i({\rm RM}\,\lambda^2 + \psi_0)},
\end{equation}
$L_{\rm 1\,GHz}$ is the flux density of the linearly polarised component at 1\,GHz, $\alpha_L$ is its spectral index, RM is the rotation measure, $\lambda = c/\nu$ is the wavelength, and $\psi_0$ is the intrinsic polarisation angle.
The RMs and spectral indices of these fits are shown in the top two rows of panels of \Fig~\ref{fig:RM}.

@@

Although there is as yet no clear consensus on whether LPTs represent one class or multiple classes of system \citep[e.g.][]{2024ApJ...961..214R}, the recent confirmation of J1912$-$4410 as a pre-evolved white dwarf pulsar \citep{2023NatAs...7..931P} and ILT\,J1101+5521 and GLEAM-X\,J0704$-$37 as polars \citep{deRuiter2025,2025A&A...695L...8R} raise the intriguing possibility that other LPTs may also be radio-emitting white dwarf pulsars at various stages of evolution \citep{2021NatAs...5..648S}.

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. J. McSweeney et al.]{
Samuel J. McSweeney$^{1}$,
Natasha Hurley-Walker$^{1}$,
Csan\'{a}d Horv\'{a}th$^{1}$,
Akash Anumarlapudi$^{2}$,
\newauthor
Angie Waszewski$^{1}$,
Dougal Dobie$^{3,4}$,
John Morgan$^{1}$,
David L.~Kaplan$^{2}$,
and Ziteng Wang$^{1}$
\\
% List of institutions
$^{1}$International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia\\
$^{4}$Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413, Milwaukee, WI 53201, USA\\
$^{2}$Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia\\
$^{3}$ARC Centre of Excellence for Gravitational Wave Discovery (OzGrav), Hawthorn, VIC 3122, Australia
}

@@

\src{} was detected blindly with the Murchison Widefield Array \citep[MWA;][]{Tingay2013} in observations taken between June--September 2024 under the `Galactic Plane Monitor', project code G0080. The prime objective of the campaign was to search for long-period transients by scanning the Galactic plane over $285^\circ < l < 65^\circ$, $|b| < 15^\circ$ on a bi-weekly cadence at 185--215\,MHz \citep[see Methods of][]{2023Natur.619..487H}; a full description will be released by Hurley-Walker et al., in prep. We also observed over June--September 2022 but made no detections in these data, as reported in \citetalias{2024MNRAS.535..909D}.
Subsequent to its redetection, we commenced a Director's Discretionary Time (DDT) campaign under project code D0042, to obtain further observations, adjusting the frequency from 200.32\,MHz to 184.96\,MHz for some of the observations to attempt to improve measurements of its DM.
Time of arrival (ToA) predictions for the DDT observations were made using a preliminary ephemeris derived from the GPM detections, which, by virtue at being at the same or similar frequencies, absorbed the delays due to dispersion and scattering into the `PEPOCH' term.

We imaged both the 2024 GPM data and the follow-up DDT data at the location of \src{} using \textsc{WSClean} \citep{2014MNRAS.444..606O}, masking \src{} itself, forming a deep model of the sky for each observation. After subtracting this from the visibilities, we phase-rotated to \src{}, and averaged the baseline data to produce a dynamic spectrum.

@@

\citetalias{2024MNRAS.535..909D} presented an exhaustive search of the ASKAP data available at the time of writing, comprising 60\,h of data, and yielding only a single detection of \src{} (in Scheduling Block ID 47253). Subsequent to the writing of \citetalias{2024MNRAS.535..909D}, a further epoch of the VAST survey was performed (SBID 63600) in which \src{} produced a detectable pulsation.

@@

The correlator data were calibrated using the standard SARAO SDP calibration pipeline, and imaged using \textsc{WSClean}. We formed a deep model of the sources within the primary beam, excepting \src{}, and subtracted this from the visibilities. We then averaged the visibilities, excluding all baselines shorter than 150\,m, to produce a dynamic spectrum of the source, shown in \Fig~\ref{fig:stacked_spectra}.

We also re-imaged the original 2023-05-28 L-band MeerKAT observation described in \citetalias{2024MNRAS.535..909D}, which was taken under project code DDT-20230525-DD-01 under capture block 1685306788. We performed a similar processing of the data, and identified four pulses with ToAs consistent with an ephemeris derived from the other data. The folded and stacked dynamic spectra are shown in \Fig~\ref{fig:stacked_spectra}.

@@

The low-frequency asymmetry is consistent with a scattering time scale of a few tens of seconds at MWA frequencies, which scales to a few tens of milliseconds at 1\,GHz.
This is in broad agreement with the model predictions of NE2001 ($\tau_{\rm sc,1GHz} = 66\,$ms) and YMW16 ($\tau_{\rm sc,1 GHz} = 75\,$ms) at the source's location assuming the DM value of 710\,pc\,cm$^{-3}$ reported in \citetalias{2024MNRAS.535..909D}.

To account for the asymmetry, we fit exponentially modified Gaussians to each pulse with the form,
\begin{equation}
  \emg(t) = A \frac{\sigma}{\tau}\sqrt{\frac{\pi}{2}}
    \exp\left(-\frac12 \left( \frac{t - \mu}{\sigma} \right)^2 \right)  \erfcx \left(\frac{1}{\sqrt{2}} \left(\frac{\sigma}{\tau} - \frac{t - \mu}{\sigma} \right) \right),
  \label{eqn:emg}
\end{equation}
where $\erfcx(X) = \exp(X^2) \erfc(X)$ is the scaled complimentary error function; $A$, $\mu$, and $\sigma$ are the height, location, and scale of an unscattered Gaussian pulse; and the scattering time scale was fixed at
\begin{equation}
      \tau(\nu) = 0.07\, \left(\frac{\nu}{\rm GHz}\right)^{-4} \, {\rm s},
\end{equation}
according to each observation's centre frequency.
A pulse stack containing all observed pulses and the corresponding unscattered pulses derived from the fits is shown in \Fig~\ref{fig:pulsestack}.

\begin{figure*}
      \centering
          \includegraphics[width=0.95\linewidth]{pulsestack.pdf}
              \caption{Pulsestack of barycentred, dedispersed profiles from the ephemeris. The dashed green vertical line marks zero phase. Baselines have been fitted and subtracted from each lightcurve. The red curves are fits of the normalised pulses to \Eqn~\ref{eqn:emg}, and the blue curves are the inferred Gaussian pulses without scattering, also normalised (see main text for details).}
                  \label{fig:pulsestack}
\end{figure*}

@@

The apparent lack of linear polarisation in the MWA pulses, however, is expected due to the high rotation measure (RM) of ${\sim}961\,{\rm rad}/{\rm m}^2$ \citepalias[as reported in][]{2024MNRAS.535..909D}, causing the pulses to be depolarised within individual channels.
We also report that no significantly circularly polarised emission was seen at MWA frequencies.

@@

The high S/N of each time bin allowed us to test whether the RM was consistent across each pulse.
We therefore performed a joint fit of the Stokes Q and U spectra of each time bin to
\begin{equation}
    Q(\nu) = {\rm Re}(\mathcal{L}(\nu)),
    \quad
    U(\nu) = {\rm Im}(\mathcal{L}(\nu)),
    \label{eqn:RM}
\end{equation}
where
\begin{equation}
    \mathcal{L}(\nu)
        = L_{\rm 1\,GHz}\left(\frac{\nu}{\rm 1\,GHz}\right)^{\alpha_L} e^{2i({\rm RM}\,\lambda^2 + \psi_0)},
\end{equation}
$L_{\rm 1\,GHz}$ is the flux density of the linearly polarised component at 1\,GHz, $\alpha_L$ is its spectral index, RM is the rotation measure, $\lambda = c/\nu$ is the wavelength, and $\psi_0$ is the intrinsic polarisation angle.
The RMs and spectral indices of these fits are shown in the top two rows of panels of \Fig~\ref{fig:RM}.

@@

Although there is as yet no clear consensus on whether LPTs represent one class or multiple classes of system \citep[e.g.][]{2024ApJ...961..214R}, the recent confirmation of J1912$-$4410 as a pre-evolved white dwarf pulsar \citep{2023NatAs...7..931P} and ILT\,J1101+5521 and GLEAM-X\,J0704$-$37 as polars \citep{deRuiter2025,2025A&A...695L...8R} raise the intriguing possibility that other LPTs may also be radio-emitting white dwarf pulsars at various stages of evolution (\citealt{2021NatAs...5..648S}, Horv\'{a}th et al. in prep).
